Optimal (Multiway) Spatial Joins

### 2024-05-26 msj.tex
yufei

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\out{\mathrm{OUT}}

@@

    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$. This means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
}
Figure~\ref{fig:ksj:cats} illustrates a tuple of each category, assuming $k=4$.

@@

\extraspacing {\bf Category 1.} Fix an arbitrary $i \in [k]$. Define $\J^\cat_i$ as the set of result tuples $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$. We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.

@@

We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
\myeqn{
    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
}
we add $(r_1, ..., r_{k-1}, r)$ to $\J^\cat_k$. Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time. The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for all $j \in [k-1]$. Hence, we have $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.

@@

Therefore, the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k^2 n \log n + k \cdot \sum_{i=1}^k |\J^\cat_i|) = O(k^2 n \log n + k \cdot \out)$.

>>>>

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\catt{\mathrm{cat2}}
\def\out{\mathrm{OUT}}

@@

    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$. This means $p$ must be the intersection point between the top edge of some rectangle $r_i$ and the left edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
}
Figure~\ref{fig:ksj:cats} illustrates a tuple of each category, assuming $k=4$.

@@

\extraspacing {\bf Category 1.} Given an $i \in [k]$, we denote by $\J^\cat_i$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$. We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.

@@

We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
\myeqn{
    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
}
we add $(r_1, ..., r_{k-1}, r)$ to $\J^\cat_k$. Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time. The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for all $j \in [k-1]$. Hence, we have $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|) = O(kn \log n + k \cdot \out)$ .

@@

Therefore, the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k) \cdot (kn \log n + k \cdot \out)$. A category-1 result tuple $\bm{t}$ may be seen more than once (this happens if the top-left corner of $B_\bm{t}$ is the top-left corner of more than one rectangle in $\bm{t}$). Duplicate removal can be achieved at no extra cost asymptotically, following the ideas explained in Section~\ref{sec:hv:type2}.

\extraspacing {\bf Category 2.} Given $i, j \in [k]$ with $i \ne j$, we denote by $\J^\catt_{i,j}$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the intersection between the top edge of $r_i$ and the left edge of $r_j$. The Category 2 of result tuples is the union of the $\J^\catt_{i,j}$ of all possible $i, j$.

top-left corner of $\bm{t}[i]$. We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
